### 2020-02-03 Common/chapters/paper_reproduction.tex
Julien and Johan: model roulette

\section{Training the neural network}
\subsection{Architecture}
The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG). This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset of 14 millions images belonging to 1000 classes. Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
It is first composed by three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks. Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function. The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit. The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax  function for classification.\\
In comparison to the original VGG, this model keeps the small filter size of 3x3 or 1x1, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions. It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one. Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/model_paper_manual.png}
\caption{Model architecture with the corresponding PyTorch code}
\label{fig:paper_model}
\end{figure}
\subsection{Criterion to save the best model}
\subsection{Tensorboard}
During the experiment the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC. These metrics are computed separately on the training and on the validation sets at the end of each epoch. They are then stored across all epochs and plotted on the same figure thanks to Tensorboard at the end of the process. Subsequently, this is also the case for the metrics measured on the test set.\\
"Tensorboard provides the visualization and tooling needed for machine learning experimentation:
\begin{itemize}
\item Tracking and visualizing metrics.
\item Visualizing the model graph (ops and layers).
\item Viewing histograms of weights, biases, or other tensors as they change over time.
\item Projecting embeddings to a lower dimensional space.
\item Displaying images, text, and audio data"\cite{39}.
\end{itemize}
In this experiment, Tensorboard is mainly used to plot the Matplotlib figures of the model performance. In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
\label{paper_tensorboard}
\subsection{Script options}
Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary. In order to achieve this goal, the Python module "argparse" is extremely useful. The latter automatically generates help/usage messages and displays errors when the argument given by the user are invalid. Table \ref{fig:paper_reproduction_options} shows all options accepted by the script.

@@

@@

\subsection{Training verification}
In deep learning, two main problems can occur during training. The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero. Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train. The second problem, the "exploding gradient" is the opposite. Concretely, in this case, large gradients are computed, which results in huge weight updates that can even reach "NaN" values. This makes the model unstable and unable to learn from training data.  \\
As a result, it is important to analyze the gradient propagation across the network to ensure that it is not facing such problems. To achieve this goal, a visualization tool to display the gradients of the entire network was implemented. This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).
\subsubsection{Gradient flow visualization}
The visualization displays the gradient flow through the entire neural network. It shows the name of each layer on the x-axis and its average gradient value on the y-axis (recall: for each layer the gradient is a matrix). Furthermore, it also gives information about the max value of the gradient and indicates in black if one of them is equal to zero.\\
The figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6. From the output layer (called "last\_layer0.weight" on the graph) to the first layer, the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.

>>>>

\section{Training the neural network}
\subsection{Architecture}
The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG). This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset of 14 millions images belonging to 1000 classes. Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
It is first composed by three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks. Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function. The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit. The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax  function for classification.\\
In comparison to the original VGG, this model keeps the small filter size of 3x3 or 1x1, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions. It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one. Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/model_paper_manual.png}
\caption{Model architecture with the corresponding PyTorch code}
\label{fig:paper_model}
\end{figure}
\subsection{Criterion to save the best model}
\subsection{Tensorboard}
During the experiment the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC. These metrics are computed separately on the training and on the validation sets at the end of each epoch. They are then stored across all epochs and plotted on the same figure thanks to Tensorboard at the end of the process. Subsequently, this is also the case for the metrics measured on the test set.\\
"Tensorboard provides the visualization and tooling needed for machine learning experimentation:
\begin{itemize}
\item Tracking and visualizing metrics.
\item Visualizing the model graph (ops and layers).
\item Viewing histograms of weights, biases, or other tensors as they change over time.
\item Projecting embeddings to a lower dimensional space.
\item Displaying images, text, and audio data"\cite{39}.
\end{itemize}
In this experiment, Tensorboard is mainly used to plot the Matplotlib figures of the model performance. In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
\label{paper_tensorboard}

\subsection{Script options}
Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary. In order to achieve this goal, the Python module "argparse" is extremely useful. The latter automatically generates help/usage messages and displays errors when the arguments type in by the user are invalid. Table \ref{fig:paper_reproduction_options} shows all options accepted by the script.

@@

\subsection{Model roulette}
As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}. Experience showed us that the exact same hyperparameters with two different initializations could give opposite results. Therefore, in order not to train the model vainly, a script which generates a given number of initializations was created. Each untrained model was then tested on the validation set since the best model during training is saved according to its performance on the validation set. The model whose score was the highest on a given metric was saved and used as base model for the experiments.

@@

\subsection{Training verification}
In deep learning, two main problems can occur during training. The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero. Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train. The second problem, the "exploding gradient" is the opposite. Concretely, in this case, large gradients are computed, which results in huge weight updates that can even reach "NaN" values. This makes the model unstable and unable to learn from training data.  \\
As a result, it is important to analyze the gradient propagation across the network to ensure that it is not facing such problems. To achieve this goal, a visualization tool to display the gradients of the entire network was implemented. This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).

\subsubsection{Gradient flow visualization}
The visualization displays the gradient flow through the entire neural network. It shows the name of each layer on the x-axis and its average gradient value on the y-axis (recall: for each layer the gradient is a matrix). Furthermore, it also gives information about the max value of the gradient and indicates in black if one of them is equal to zero.\\
The figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6. From the output layer (called "last\_layer0.weight" on the graph) to the first layer, the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.
